---
title: 'Tutorial 6: Linear Regression (Part 2 Deadline: 16 Oct 2023 9am)'
author: 'REPLACE WITH YOUR NAME and ID#'
date: 'Due by 16 Oct 2023, 9:00 AM'
output: html_document
---
## Preparation

```{r load-libraries, echo=TRUE, message=FALSE, warning=FALSE}
# load required packages
library("tidyverse")
library("readxl")
library("rpivotTable")
library("knitr")
library("kableExtra")
library("RColorBrewer")
library("psych")
library("ggplot2")
```

## Part One: Lab-Session Completion and Discussion

-   Dataset required: `WorldBankData.csv`

(Note: This dataset comes from a publically available dataset from The
World Bank.
<https://databank.worldbank.org/source/world-development-indicators>.)

First, load in the dataset for this question. There are 8 variables in
this real-world dataset, from 258 countries in 2016/2017: -
Country.Name.Name: name of country - Country.Code: code given to
country - `Human.Capital.Index`: unitless number that goes from 0
to 1. - `GDP.per.capita.PPP` in US Dollar. This is GDP per capita, but
taking into account the purchasing power of the local currency, by
comparing how much it costs to buy a basket of goods (e.g. food)
compared to the reference currency (USD). (PPP stands for Purchasing
Power Parity) - `Health.Expenditure.per.capita` in US Dollar. -
`Tertiary.Education.Expenditure.per.student` in US Dollar. -
`Population`. - `Life.Expectancy.at.birth` in years. -
`Diabetes.Prevalence` in units of % of population ages 20 to 79. -
`Years.of.Compulsory.Education` in years.

Being a data set in real world, there are lots of missing data. Be wary
of this!

```{r q1-read-dataset, echo=TRUE}

dta_wb = read.csv('WorldBankData.csv')

```

### Question 1

First, let's investigate `Human.Capital.Index`. As noted by Prime
Minister Lee in his 2019 National Day Rally, Singapore topped the world
on this Human Capital Index in 2018. Let's try to see what are some of
the possible variables that correlate with this.

(1a) Start off by plotting `Human.Capital.Index` (on the y-axis) versus
`GDP.per.capita.PPP` on the x-axis. Why are we constructing this plot?
What do you notice? What type of variable relationship exists? Is it
linear?

```{r 1a), echo=TRUE}
plot(dta_wb$GDP.per.capita.PPP , dta_wb$Human.Capital.Index)

```

::: {style="color: red"}

:::

(1b) What type of transformation could you apply? Try a few functions
that were shown in class: `x^2, x^3, ...`, `exp(x)`, `log10(x)`. Make a
plot that shows a linear relationship, and describe what you did.

For fun: add code into your plot to highlight the dot that represents
Singapore.

```{r 1b), echo=TRUE}
dta_wb$pw3GDP = (dta_wb$GDP.per.capita.PPP)^3
plot(dta_wb$pw3GDP , dta_wb$Human.Capital.Index, col = ifelse(dta_wb$Country.Name == "Singapore", "red","black"))

dta_wb$pw2GDP = (dta_wb$GDP.per.capita.PPP)^2
plot(dta_wb$pw2GDP , dta_wb$Human.Capital.Index, col = ifelse(dta_wb$Country.Name == "Singapore", "red","black"))

dta_wb$logGDP = log10(dta_wb$GDP.per.capita.PPP)
plot(dta_wb$logGDP , dta_wb$Human.Capital.Index, col = ifelse(dta_wb$Country.Name == "Singapore", "red","black"))

```
::: {style="color: red"}
**Type your answer here**
:::

(1c) Now that you have a plot of a linear relationship, run a linear
regression using `lm()`, predicting `Human Capital Index`. Run
`summary()` on the `lm` object to produce an output table. Interpret the
output of the `lm()`. What do the `b` coefficients mean? (Interpret them
"in English" and try to make sense of the numbers, even if they might
seem weird at first. How many countries made it into this regression?
(What happened to the rest?) Comment on the goodness-of-fit statistics.

```{r 1c), echo=TRUE}
summary(lm(Human.Capital.Index ~ logGDP, dta_wb))
```
::: {style="color: red"}
**Type your answer here**
:::

### Question 2

-   Dataset required: `WorldBankData.csv`

Let's look at another set of variables in the same dataset. This time,
let's consider `Health.Expenditure.per.capita`, `Diabetes.Prevalence`,
and `Life.Expectancy.at.birth`.

(2a) If you had to design a predictive hypothesis with these three
variables, what would it be? Which would be your dependent variable, and
which would be your independent variables? Justify your answer. (Note,
there is no necessarily "right" or "wrong" answer for this question, as
is the case in real life, but there are more justifiable answers that
you would feel more comfortable putting up to your boss!)


plot(dta_wb$GDP.per.capita.PPP , dta_wb$Human.Capital.Index)

::: {style="color: red"}
**Type your answer here**
:::

(2b) Plot the bivariate relationships between these three variables. (In
other words, plot x-y scatterplots. There are 3 variables, so you'll
need 3 scatterplots.) Please also apply the same transformation in (1b)
to `Health.Expenditure.per.capita`. Comment on the relationship between
the variables.


```{r 2b), echo=TRUE}
plot(dta_wb$Life.Expectancy.at.birth , log10(dta_wb$Health.Expenditure.per.capita))
plot(dta_wb$Life.Expectancy.at.birth , dta_wb$Diabetes.Prevalence)
plot(log10(dta_wb$Health.Expenditure.per.capita), dta_wb$Diabetes.Prevalence)
```

::: {style="color: red"}
**Type your answer here**
:::

(2c) Run a multiple regression predicting `Life.Expectancy.at.birth`
using the other two variables. Interpret the coefficients, spelling out
what the numbers mean. Comment on your answers.

```{r 2c), echo=TRUE}
summary(lm(Life.Expectancy.at.birth ~ log10(Health.Expenditure.per.capita) + Diabetes.Prevalence, dta_wb))
```

::: {style="color: red"}
**Type your answer here**
:::
